### 2020-08-05 README.Rmd
initial structure

---
title: "Donations' tax in Chile"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Structure of the repo

This repo contains the code of the article entitled _More flexibility reduces social donations_, the main files are:

- [article-elsevier.Rmd](article-elsevier/article-elsevier.Rmd): generates the manuscript
  - [mybibfile.bib](article-elsevier/mybibfile.bib): contains the database of references
- [presentation.Rmd](presentation/presentation.Rmd): reproduces a presentation of the article