### 2020-08-05 article-elsevier/article-elsevier.Rmd

---
title: More flexibility reduces social donations
author:
  - name: Sebastián Cea-Echenique
    email: sebastian.cea@pucv.cl
    affiliation: School of Industrial Engineering, Pontificia Universidad Católica de Valparaíso
    footnote: 1
  - name: Claudia Toledo
    email: claudia@example.com
    affiliation: School of Industrial Engineering, Pontificia Universidad Católica de Valparaíso
#   - name: Cat Memes
#     email: cat@example.com
#     affiliation: Another University
#     footnote: 2
#   - name: Derek Zoolander
#     email: derek@example.com
#     affiliation: Some Institute of Technology
#     footnote: 2
# address:
#   - code: Some Institute of Technology
#     address: Department, Street, City, State, Zip
#   - code: Another University
#     address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  # - code: 2
    # text: "Equal contribution"
abstract: |
  We exploit a change in tax donations' law to evaluate a change to more flexibility.

journal: "Economics Letters or Journal of Public Economics"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

<!-- _Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_ -->

Introduction
==========================

#### Motivation and legal background

Law XX introduced a framework to foster social donations. Several changes to that law were made...

#### Insertion in the Literature

The relation between taxes and social donations has been studied in the recent literature by ...

Empirical strategy
============

We use a differences in differences approach in a panel structure of the data.

References {#references .unnumbered}
==========