### 2022-05-10 article-elsevier/article-elsevier.tex
permisos

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Economics Letters or Journal of Public Economics} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={More flexibility reduces social donations},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{More flexibility reduces social donations}
    \author[School of Industrial Engineering, Pontificia Universidad Católica de
Valparaíso]{Sebastián Cea-Echenique\corref{1}}
   \ead{sebastian.cea@pucv.cl} 
    \author[School of Industrial Engineering, Pontificia Universidad Católica de
Valparaíso]{Claudia Toledo}
   \ead{claudia@example.com} 
        \cortext[1]{Corresponding Author}

  \begin{abstract}
  We exploit a change in tax donations' law to evaluate a change to more
  flexibility.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{motivation-and-legal-background}{%
\paragraph{Motivation and legal
background}\label{motivation-and-legal-background}}

Law XX introduced a framework to foster social donations. Several
changes to that law were made\ldots{}

\hypertarget{insertion-in-the-literature}{%
\paragraph{Insertion in the
Literature}\label{insertion-in-the-literature}}

The relation between taxes and social donations has been studied in the
recent literature by \ldots{}

\hypertarget{empirical-strategy}{%
\section{Empirical strategy}\label{empirical-strategy}}

We use a differences in differences approach in a panel structure of the
data.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\end{document}